A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities

### 2012-05-21 papers/contact/paper.tex
Some changes to paper, some data for walls .1 fillfrac that has been computed since bug fix Tested everything on bennet.

When we consider an inhomogeneous system, we may prefer to ask for
\emph{which} sphere we are interested in computing the contact
density.  One way to answer this is to specify a sphere at a fixed
position $\mathbf{r}$.  By making the hard-sphere radius dependent on
position, we can compute the pressure on the hard spheres located at
$\mathbf{r}$.
\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^A}\\
  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
  g_\sigma^A(\rr) &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
Equation~\ref{eq:g-A-exact} is an exact expression for $g_\sigma^A$
given the exact functional $A_{HS}$.  In practice, we use the
approximate White Bear functional, and find an approximate expression
for $g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
expression will involve more convolutions than are present in the free
energy itself, making this formulation relatively expensive.

@@

@@

An alternate definition of the averaged correlation function for an
inhomogeneous system would be the correlation function averaged for
spheres \emph{touching at a given point}.  This would be a symmetric
measure, since both spheres involved are treated in the same way.  In
fact, the density of such spheres is just the density $n_0(\rr)$ from
FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
is the fundamental measure of sphere surface contact, it makes sense
to define a corresponding correlation function $g_\sigma^S$.  This is
\emph{almost} what is done by Yu and  Wu

\fixme{Analogous to the assymetric case is the symmetric, for which 
we are no longer able to simply pick a hypothetical sphere centered 
at position $r$ and expand its radius.  Instead we must formulate a way 
to expand the radii of spheres that just touch a specific point of 
contact and only expand them at that point of contact, in a symmetric 
fashion.  Therefore, when we compute the neccasary derivitives wih 
respect to $R$', we expand a multitude of spheres, all of them positioned 
around and just touching the point $\mathbf{r}$.  When their radii 
expand, each does not expand outward in all directions as in the 
case of A, rather they each expand outward at only the point $\mathbf{r}$.  One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains 
at radius $R$.  When we expand each sphere that touches $\mathbf{r}$ 
in this peculiar fashion, we have in the end the same amount of shell 
area protruding outward as we would if we had simply expanded one 
sphere in fashion of case A.  Also the protrusions, once added up 
for all spheres sourrounding $\mathbf{r}$, will yield a symmetry 
in their direction of protrusion.  Each protruding sphere bit will 
expand outward from its center and all directions will be covered uniformly.
See Figure \ref{fig:contact}.

@@

These qualitative concepts lead us to derivations for the $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ term in our derivation of the pressure, eq \ref{eq:p_{HS}}.  The rest of the derivation results when we recognize that instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ as in the assymmetric case, we will instead divide by $n_0(\mathbf{r})^2$ since we have symmetry around the point $\mathbf{r}$.  This will yield our symmetric correlation, $g_\sigma^S$:

\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
  g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
\end{align}}Complete expressions for $g_\sigma^S$ are included in
Appendix~\ref{appendix:g-S}.\\

>>>>

When we consider an inhomogeneous system, we may prefer to ask for
\emph{which} sphere we are interested in computing the contact
density.  One way to answer this is to specify a sphere at a fixed
position $\mathbf{r}$.  By making the hard-sphere radius dependent on
position, we can compute the pressure on the hard spheres located at
$\mathbf{r}$.
\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^A}\\
  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \label{eq:g-A-exact-a}\\
  g_\sigma^A(\rr) &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
Equation~\ref{eq:g-A-exact} is an exact expression for $g_\sigma^A$
given the exact functional $A_{HS}$.  In practice, we use the
approximate White Bear functional, and find an approximate expression
for $g_\sigma^A$, which is given in the Appendix.  This
expression will involve more convolutions than are present in the free
energy itself, making this formulation relatively expensive.

In addition to the different densities in the denominators of Equations 
\ref{eq:g-A-exact-a} and \ref{eq:g-S-exact-a} there is also a difference 
between the treatment of the A and S cases in the derivatives 
$\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
that we use involve a number of weighted densities and for the derivatives 
of these we use two different conceptual approaches.  In the assymetric case, 
a change in $R(\mathbf{r})$ will refer to a change in the radius of a sphere 
that is located at the position $\mathbf{r}$.  The weighted densities that are 
affected by such a change will be those that are located a distance $R$ away 
from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.

@@

For the computation of the symmetrically averaged correlation function 
we must formulate derivatives with respect to $R$ in a symmetric fashion.  
Conceptually we expand the radii of spheres that just touch a specific 
point of contact and only expand them outward at that point of contact.  
Therefore $R(\mathbf{r})$ in our calculations refers to the radii of 
spheres that touch $\mathbf{r}$ and $\delta R(\mathbf{r})$ refers to a 
change of these radii at only the point $\mathbf{r}$.  One can imagine 
a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell 
area protruding outward at $\mathbf{r}$.  Figure~\ref{fig:contact} 
illustrates such a change in $R$.  The full derivations are found in Appendix~\ref{appendix:g-S}

These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$, 
which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the 
derivation of the correlatation function results when we recognize that 
instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ 
as in the assymmetric case we will instead divide by $n_0(\mathbf{r})^2$ 
since we have symmetry around the point $\mathbf{r}$.

\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \label{eq:g-S-exact-a}\\
  g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-S-exact}
\end{align}Complete expressions for $g_\sigma^S$ are included in
Appendix~\ref{appendix:g-S}.\\

@@

@@
